---
title: " week tutorial"
author: "by: René van Son"
bibliography: references.bib
---
# Preliminary Research Question and Literature

## Research question
For my research on collaboration networks, I am (for now) interested in the following question: 
"*Does the development of an early-carreer interdisciplinary network (i.e. a co-authorship focused more on collaboration across faculties/departments) lead to the publication of high-impact work?*"

On the macro level, this research would *describe* clustering in collaboration networks; on the micro-level, this research will *explain* how this clustering is due to differences in early-stage interdisciplinary collaborations.

I found the following relevant articles 

- Haythornthwaite, C. (2006). Learning and knowledge networks in interdisciplinary collaborations. Journal of the American Society for Information Science and Technology, 57(8), 1079–1092. (https://doi.org/10.1002/asi.20371) 

- van Rijnsoever, F. J., & Hessels, L. K. (2011). Factors associated with disciplinary and interdisciplinary research collaboration. Research Policy, 40(3), 463–472. https://doi.org/10.1016/j.respol.2010.11.001

These articles deal with the different aspects of social networks in research collaboration. Haythornthwaite (2006) argues that knowledge exchange is particularly high in interdisciplinary research teams. Van Rijnsoever and Hessels' (2011) work centers around the different characteristics of researchers that could lead to them engaging in either interdisciplinary or non-interdisciplinary collaborations.   



## R-Tutorial
```{r, echo=FALSE}
########################### Title: René van Son - 04-09-2024#################

# start with clean workspace
rm(list = ls())
```


```{r, echo=FALSE, eval=FALSE}
# install.packages I will need later here
#install.packages("installr")  # you  first install packages
require(installr)  # then you will need to load them. This package is used to simply update R
#install.packages("foreign")
#install.packages("tidyverse")

```


```{r}
require(foreign)  # used to read in spss data files
require(tidyverse)
# update if necessarry. Best to run this command in RGui, not in RStudio.
#updateR()
```


```{r, echo=FALSE}
# define workdirectory, note the double backslashes
setwd("/Users/jannevanheesch/Documents/GitHub/labjournalRVS")  # change to your own workdirectory
getwd()#check

testvariable <- 4
ls()
list <- ls()

# ignore the warnings ?read.spss note that I have saved the data files in a folder called
# 'addfiles'.

cv08 <- foreign::read.spss("tutorial_data/Cultural_Changes_2008.sav", use.value.labels = T, to.data.frame = T)
cv10 <- foreign::read.spss("tutorial_data/Cultural_Changes_2010.sav", use.value.labels = T, to.data.frame = T)

# normally I think setting use.value.labels=F is more convenient. Thus lets load the data again but
# now without labels
cv08_nolab <- foreign::read.spss("tutorial_data/Cultural_Changes_2008.sav", use.value.labels = F, to.data.frame = T)
cv10_nolab <- foreign::read.spss("tutorial_data/Cultural_Changes_2010.sav", use.value.labels = F, to.data.frame = T)

# finally, import the data using haven
cv08_haven <- haven::read_spss("tutorial_data/Cultural_Changes_2008.sav")
cv10_haven <- haven::read_spss("tutorial_data/Cultural_Changes_2010.sav")

str(cv08) 
str(cv08_haven)
str(cv08_nolab)

str(cv08$lftop) 
str(cv08_haven$lftop)
str(cv08_nolab$lftop)

attributes(cv08$lftop)
attributes(cv08_nolab$lftop)
attributes(cv08_haven$lftop)

table(cv08_haven$lftop, useNA = "always")

names(cv08_haven)
summary(cv08_haven)
head(cv08_haven)
# fix(cv08_haven) #will produce an error
#fix(cv08)
#View(cv08_haven)

#create new variable from variable lftop
cv08$lftop_new <- cv08$lftop

#now, assign 'onbekend' as missing
cv08$lftop_new[cv08$lftop == "Onbekend"] <- NA

#so, still not NA? but 
table(cv08$lftop_new, useNA = "always")
levels(cv08$lftop_new)

#so check
str(cv08$lftop_new)

cv08$agen <- as.numeric(as.character(cv08$lftop_new))  # how clumsy. we first convert the factor to a string and then to a numeric variable.

#check
table(cv08$agen, useNA = "always")

#you can also use mutate to create a new age variable, with 99 as missing value. This is faster, but needs tidyverse.
cv08_haven <- mutate(cv08_haven, lftop_new = na_if(lftop, 99))

levels(cv08$var006n)
table(cv08$var006n, useNA = "always")

# lets make it a numeric var first
cv08$educn <- as.numeric(cv08$var006n)
# check
table(cv08$educn, useNA = "always")

#start with empty variable (everything missing)
cv08$educ3 <- NA

cv08$educ3[cv08$educn == 2 | cv08$educn == 3] <- 1
cv08$educ3[cv08$educn > 3 & cv08$educn < 8] <- 2
cv08$educ3[cv08$educn > 7 & cv08$educn < 11] <- 3

table(cv08$educ3, useNA = "always")
prop.table(table(cv08$educ3, useNA = "always"))

cv08$educ3 <- as.factor(cv08$educ3)
table(cv08$educ3, useNA = "always")

levels(cv08$educ3) <- c("primary", "secondary", "tertiary")
table(cv08$educ3, useNA = "always")

#install.packages("labelled")
require(labelled)  # to be able to use the recode function on haven labelled variables

# inspect variable
str(cv08_haven$var006n)

attr(cv08_haven$var006n, "labels")
table(cv08_haven$var006n)

cv08_haven <- cv08_haven %>%
    mutate(educ3 = recode(var006n, `-3` = -9, `-1` = 1, `1` = 1, `2` = 2, `3` = 2, `4` = 2, `5` = 2,
        `6` = 3, `7` = 3, `8` = 3, `10` = -9, .keep_value_labels = FALSE), educ3 = na_if(educ3, -9),
        educ3 = factor(educ3, levels = c(1, 2, 3), labels = c("primary", "secondary", "tertiary")))
table(cv08_haven$educ3)

summary(cv08$int055)
summary(cv08$int056)
summary(cv08$int057)

#make numeric instead of factors
cv08$int055n <- as.numeric(cv08$int055)
table(cv08$int055n, useNA = "always")
cv08$int056n <- as.numeric(cv08$int056)
table(cv08$int056n, useNA = "always")
cv08$int057n <- as.numeric(cv08$int057)
table(cv08$int057n, useNA = "always")

#Assign missing values (everything under 5 = missing, because 3='don't know')
cv08$int055n[cv08$int055n < 5] <- NA
#now it makes more sene to reassign values, where 5=1, 6=2, etc.
cv08$int055n <- cv08$int055n - 4

#repeat
cv08$int056n[cv08$int056n < 5] <- NA
cv08$int056n <- cv08$int056n - 4
cv08$int057n[cv08$int057n < 5] <- NA
cv08$int057n <- cv08$int057n - 4

# How does the function mean work in R?
mean(cv08$int055n)  # whoops
mean(cv08$int055n, na.rm = TRUE)  # works.  but not what we want.
mean(c(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)  # works but not what we want.


testmeans <- rowMeans(cbind(cv08$int055n, cv08$int056n, cv08$int057n), na.rm = T)
head(testmeans)  # yes!

# lets first count how many missings we have for each respondent
nmis <- rowSums(is.na(cbind(cv08$int055n, cv08$int056n, cv08$int057n)))

#if the number of missings is two or higher, we do not want a mean
testmeans <- ifelse(nmis < 2, testmeans, NA)

# add the calculated means to our dataset
cv08$int_mean <- testmeans







```




```{r}
#How to merge datasets. You want to select some variables (in this case: sex, education, health, region, age, and id). For both 08 and 10.
cv08_sel <- cv08[, c("we_id", "lftop", "geslacht", "var006n", "v401", "landd")]
cv10_sel <- cv08[, c("we_id", "lftop", "geslacht", "var006n", "v401", "landd")]

#make same variable names 
names(cv08_sel) <- names(cv10_sel) <- c("id", "age", "sex", "educ", "health", "region")

#check if id is factor
levels(cv08_sel$id)
#
attributes(cv08_sel$id) #yes it is

#change to numeric, again with the weird method of first transforming it to string
cv08_sel$id <- as.numeric(as.character(cv08_sel$id))
cv10_sel$id <- as.numeric(as.character(cv10_sel$id))

#check
table(cv08_sel$id, useNA = "always") #yes

# do the same for age
attributes(cv08_sel$age)

cv08_sel$age <- as.numeric(as.character(cv08_sel$age))
cv10_sel$age <- as.numeric(as.character(cv10_sel$age))

attributes(cv08_sel$age)

#now for gender (so make dummy)
levels(cv08_sel$sex)
levels(cv10_sel$sex)
table(cv08_sel$sex, useNA = "always")
table(cv10_sel$sex, useNA = "always")
attributes(cv08_sel$sex) #class=factor, so make numeric

cv08_sel$sexn <- as.numeric(cv08_sel$sex)
table(cv08_sel$sexn)

# recode into dummy. So: R checks if value for sex =2, then the new value=1. All other old values get 0. 
cv08_sel$men <- ifelse(cv08_sel$sexn == 2, 1, 0)
cv08_sel$men <- ifelse(cv08_sel$sexn == 1, NA, cv08_sel$men)

# check
table(cv08_sel$men, useNA = "always")


```



