
 QUESTIONABLE RANGE, CONSIDER CHANGING
#Okay, now we have all the works by Radboud sociologists in two waves (2014-2019, 2020-2024). Now we must find some way to add information about how interdisciplinary these works are. One way would be to look at the 'related works', provided by open alex. Another way would be to look at works cited. Then it is necessary to look where these works come from (i.e. their discipline? or their topics/concepts?)

```{r}
rm(list = ls())

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```

```{r}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
#devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```

RSiena models evolution of network. So, observations at T1 is given. What would need to happen to come at observations at T2. These are the 'rules' (e.g. if we have the same gender, the possibility is so much higher that you befriend someone with the same gender). So: interpretations of all effects are based/conditional on the network we observe at T1. So T1 compositions of networks is really important. 

If you *know* the rules of these agents, we could simulate how networks evolve. So, we **assume** the rules (reciprocity, gender homophily, etc.). In SNA, the statistics (i.e. the rules) are then tweaked to get to T2 in the best way. 

Rsiena assumes that netowrk changes occur by ministeps. 

```{r}
ts_net1

net1g <- graph_from_adjacency_matrix(ts_net1, mode = "directed")
coords <- layout_(net1g, nicely())  #let us keep the layout
par(mar = c(0.1, 0.1, 0.1, 0.1))
{
    plot.igraph(net1g, layout = coords)
    graphics::box()
}

#to...
set.seed(24553253)
ego <- ts_select(net = ts_net1, steps = 1)  #in rsienatwostep two actors may make a change together but here not
ego

options <- ts_alternatives_ministep(net = ts_net1, ego = ego)
#options
plots <- lapply(options, graph_from_adjacency_matrix, mode = "directed")
par(mar = c(0, 0, 0, 0) + 0.1)
par(mfrow = c(1, 2))

fplot <- function(x) {
    plot.igraph(x, layout = coords, margin = 0)
    graphics::box()
}

lapply(plots, fplot)
```

Degrees. You can check in the plots.
```{r}
ts_degree(net = options[[1]], ego = ego)
# or for all options
lapply(options, ts_degree, ego = ego)
```
Reciprocity. 
```{r}
lapply(options, ts_recip, ego = ego)
```

Evaluation function. This is: summing over K (the statistics). The Betas are the 'importance scores'
```{r}
option <- 4
ts_degree(options[[option]], ego = ego) * -1 + ts_recip(options[[option]], ego = ego) * 1.5
```
Eval-function does this for you
```{r}
eval <- ts_eval(net = options[[option]], ego = ego, statistics = list(ts_degree, ts_recip), parameters = c(-1,
    1.5))
eval
```

```{r}
eval <- sapply(options, FUN = ts_eval, ego = ego, statistics = list(ts_degree, ts_recip), parameters = c(-1,
    1.5))
eval
print("network with maximum evaluation score:")
which.max(eval)
```

These evaluation scores 
```{r}
choice <- sample(1:length(eval), size = 1, prob = exp(eval)/sum(exp(eval)))
print("choice:")
choice
#print('network:') options[[choice]]
```


Okay let's simulate with two statistics (degree=-1, and recip=1.5) and rate 2.
```{r}
#rate <- 2
#degree <- -1
#recip <- 1.5
# we do not calculate SE for now.
#ans <- ts_estim(net1 = s501, net2 = s502, statistics = list(ts_degree, ts_recip), p2step = c(1, 0, 0), conv = 0.01, phase3 = FALSE, itef3 = 50, verbose = TRUE, startvalues = c(5,2,2)) #does not work for now, but it spits out parameter values
```

```{r}
#rate point at the average amount of ministeps each individual has to take. The simplest interpretation of the other parameters are similar to logit models. I.e., a positive parameter  value is interpreted as a 'positive' effect'. RSiena only tests: for egos, given the network at T1, certain statistics are more important than other statistics. So:
#1. interpretations are dependent at T1
#2. Rsiena is on ego level


```

