# Webscraping in R
```{r}
# start with clean workspace
rm(list = ls())
#install.packages('data.table')
library(data.table)  # mainly for faster data handling
#library(tidyverse)  # I assume you already installed this one!
#install.packages('httr') 
# we don't need this for now require(httr)
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
#install.packages("devtools")
require(devtools)
# Note we're doing something different here. We're installing a *latest* version directly from
# GitHub This is because the released version of this packages contains some errors!
#devtools::install_github("jkeirstead/scholar")

require(scholar)

# define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'

```
Now the packages for webscraping are installed.

```{r}
soc_staff <- read_html("https://web.archive.org/web/20230528153336/https://www.ru.nl/sociology/research/staff/")
```
But what is this?

```{r}
class(soc_staff)
```


The whole website is not neccessary, so we need to 'filter' 
```{r, eval=FALSE}
soc_staff <- soc_staff %>%
    rvest::html_element("body") %>%
   rvest::html_elements("td") %>%
    rvest::html_text()

soc_staff
```
But can we scrape load the table at once?
```{r}
soc_staff <- soc_staff %>%
    rvest::html_element("body") %>%
  rvest::html_table()
```
Try the same for Leiden
```{r}
pol_staff <- read_html("https://www.universiteitleiden.nl/en/social-behavioural-sciences/political-science/staff#tab-1")
```

Now, 'filter' again
```{r}
pol_staff2 <- pol_staff %>%
    rvest::html_element("body") %>%
   rvest::html_elements("ul.table-list") %>%
   rvest::html_elements("li") %>%
    rvest::html_elements("a") %>%
    html_elements("div") %>%
    html_elements("strong") %>%
    html_text()
   
pol_staff2
```

```{r}
#install.packages("RSelenium")
require(RSelenium)
```

You can also look for authors in Open Alex. For example:
```{r}
Jochem_URL <- read_html("https://api.openalex.org/authors?q=jochem+tolsma") %>% html_text2()
Jochem_URL
substr(Jochem_URL, 1, 100)
```
But this is a bit of an annoying file. JSON.
```{r}
#install.packages("openalexR")
require(openalexR)
require(dplyr)
```

```{r}
#install.packages("jsonlite")
require(jsonlite)
jt_json <- fromJSON("https://api.openalex.org/authors?search=Jochem+Tolsma", simplifyVector = FALSE)
glimpse(jt_json, max.level = 1) 
jt_json[["results"]][[1]][["display_name"]]

Fetch_Jochem <- oa_fetch(entity = "author", search = "Jochem Tolsma")
Fetch_Jochem

```
We have the object Fetch_Jochem, which contains the information about Jochem Tolsma. Now let's try to look for institutions.
```{r}
RU_id <- oa_fetch(entity = "institutions", search = "Radboud University")$id[1] 
UvA_id <- oa_fetch(entity = "institutions", search = "University of Amsterdam")$id[1] 

```
We get 2 ids, one for UMC, one for RU. 

Let's create a network (adjacency matrix) from these 4 authors:

Jochem Tolsma
Tom van der Meer
Maurice Gesthuizen
Michael Savelkoul

```{r}
TvdM <- oa_fetch(entity = "author", search = "Tom van der Meer", affiliations.institution.id = UvA_id)
MG <- oa_fetch(entity = "author", search = "Maurice Gesthuizen")
MS <- oa_fetch(entity = "author", search = "Michael Savelkoul")

TvdM
MG
MS


TvdM_works <- oa_fetch(entity = "works", author.id = TvdM$id[1])
MG_works <- oa_fetch(entity = "works", author.id = MG$id)
MS_works <- oa_fetch(entity = "works", author.id = MS$id)
JT_works <- oa_fetch(entity = "works", author.id = Fetch_Jochem$id)

TvdM_works
MG_works
MS_works
JT_works


```
Let's do two waves of publications and make an adjacency matrix from there. 2014-2018 and 2019-2023.



```{r}
Wave1 <- matrix(c(0,0,0,0, 
                   0,0,0,0,
                   0,0,0,0,
                   0,0,0,0),nrow = 4, ncol = 4,byrow = TRUE)
Wave1

rownames(Wave1) <- c(TvdM$id[1], MG$id, MS$id, Fetch_Jochem$id)
colnames(Wave1) <- c(TvdM$id[1], MG$id, MS$id, Fetch_Jochem$id)
```

```{r}
Wave2 <- matrix(c(0,0,0,0, 
                   0,0,0,0,
                   0,0,0,0,
                   0,0,0,0),nrow = 4, ncol = 4,byrow = TRUE)
rownames(Wave2) <- c(TvdM$id[1], MG$id, MS$id, Fetch_Jochem$id)
colnames(Wave2) <- c(TvdM$id[1], MG$id, MS$id, Fetch_Jochem$id)
```


Now loop
```{r}
works <- rbind(TvdM_works, MG_works, MS_works, JT_works)
works$dups <- duplicated(works)
sum(works$dups)
works[1,]$author

works <- works %>% filter(!duplicated(id))
ego <- works[1,]$author[[1]]$au_id[1]
alter <- works[1,]$author[[1]]$au_id[-1]

Wave2[ego, alter] <- 1
Wave2
```
But 
