---
editor_options:
  markdown:
    wrap: 72
---

# Elaboration on Introduction and Network Visualization

For my research on interdisciplinarity in collaboration networks, it is
useful to proceed stepwise. For now, I think I need the following steps:

My descriptive research question is:

*RQ1: did collaboration networks become more interdisciplinary over
time, i.e. did the structure of the collaboration network change over
time with regards to the frequency/proportion of interdisciplinary
collaborations?*

1.  I think this means that I should gather all collaboration partners
    of all sociology staff-members of the RU at t1.

<!-- -->

2.  Then, I need to gather all collaboration partners of all sociology
    staff-members of the RU at t2.

<!-- -->

3.  Then, I need *some* measure of interdisciplinarity as network
    characteristic (e.g.\@luvzar2014community).

My explanatory research question is:

*RQ2: to what extent does increasing interdisciplinarity at the
macro-level(=network) lead to selective/partial clustering at the micro
level?*

So, this question is about how structural factors lead to clustering
around some individuals, or how network characteristics may lead to some
individuals acting as collaboration-brokers. Or: are there
selection-effects of researchers who are interdisciplinary-focused?

4.  To engage with this question, I think it is necessary to gather
    relationship characteristics. So, after determining the degree of
    interdisciplinarity of the network, I could look at the ego-net
    level. I expect that interdisciplinary collaborations on the
    macro-level have a kind of mushrooming effect on relations on the
    micro-level, as interdisciplinary research collaborations
    automatically mean new connections (at departments, faculties,
    etc.). Following this I could, for example, look at the consequences
    (in terms of homophily, for example) if the network indeed turns out
    to be clustered around interdisciplinarity. This would result in the
    observation that researchers who engage in interdisciplinary
    research a lot are more likely to engage in *even more*
    interdisciplinary research, and researchers who do not regularly
    engage in interdisciplinary research become more 'isolated'. This
    would result in less in- and out-degrees. DOES THIS MAKE SENSE???

NOTES: look at openAlex at how their logic works with disciplinarity.
\## Notes Week 4

We did not yet cover whole network characteristics. Of course,
clustering in ego-nets will differ from socio-nets. We lso talked about
selection and influence. This is a bit different in a whole socionet.
Because here, one is connected with people across different distances.

FOR MY PURPOSE, HOW DO I MEASURE INTERDISCIPLINARITY? YOU COULD MEASURE
IT AS 1/0, INTERDISCIPLINARITY OR NOT, BUT THEN YOU LOSE ALOT OF
INFORMATION. SO, THERE SHOULD BE A MORE CONCISE WAY. YOU CAN ALSO LOOK
AT OVERLAP AND DISTANCE BETWEEN DISCIPLINES. E.G. CRIMINOLOGY AND
INCLUSION/EXCLUSION PROCESS. HOW ARE THESE THEMES LINKED? ONE OPTION IS
TO FIRST CREATE DIVERSITY INDEX (PROPORTION OF RESEARCH THAT A
RESEARCHER HAS DONE IN POLITICAL SCIENCE, SOCIOLOGY), ETC. AND THEN
CREATE A WEIGHT TO IT. E.G. WHAT IS THE CHANCE THAT ONE IS A
SOCIOLOGIST, ONE IS A POLITICAL SCIENTIST, BUT TAKING INTO ACCOUNT THE
RELATION *BETWEEN* THESE DISCIPLINES. E.G. SOCIOLOGY AND POLSCI IS
CLOSER THAN SOCIOLOGY AND NATURAL SCIENCES.

1.  SO FIRST GIVE PROPORTION TO RESEARCHER
2.  THEN CREATE SOME DISTANCE SCORE BETWEEN DISCIPLINES
3.  THEN CREATE SOME DISTANCE SCORE BETWEEN SCIENTIST

This relates to segregation. How do we 'calculate' segregation? In this
case, node-attributes should hang together. For example, the product of
all node-attributes who are the same:

Sigma_ij(X_i \* X_j \* w_ij)

Standardized: Sigma_ij(z_i \* X_j \* w_ij)

But, control for network size. Maybe:

(Sigma_ij(z_i \* X_j \* w_ij))/N

But, we need to take into account all ties

(Sigma_ij(z_i \* X_j \* w_ij))/(N(w_ij))

But above, w_ij is defined as: is there a relation 1/0. But some nodes
are even further away from other nodes. So, w_ij could also be defined
as where the distance is AT LEAST 2, but that gives other segregation
score! So, w_ij is a distance score. It is very important how you define
it.

So now we know that we can have a segregation measure. Now back to
influence. Are nodes influenced equally strong by their ego-network
*regardless how large that network is* (row standardize), or do all the
alters have the same influence on i (not row standardize). In other
words, how do you theoretically define influence-processes in a network.




## Exercise

```{r echo=TRUE, warning=FALSE, message=FALSE}
#load R package to plot
require(igraph)
```

```{r}
#Plot graph
g <- make_graph("Zachary")
plot(g)
```

Okay, let's show the adjacency matrix that belongs to it.

```{r}
gmatrix <- as_adjacency_matrix(g, type="both", sparse=FALSE)
gmatrix
```

### Descriptives

Okay, let's explore

```{r}
#to get the number of nodes
vcount(g)
#to count number of edges, ties
ecount(g)
```

#### Degree

Okay, so we have 34 nodes and a total of 78 ties. Let's look at degree.
What was it again?

```{r}
igraph::degree(g)
dyad.census(g)
```

This is the amount of indegrees of each node. But let's take another
look. In the histogram we can see that there are many people with a few
indegrees, but not many people with a lot of indegrees.

```{r}
hist(table(degree(g)), xlab='indegree', main= 'Histogram of indegree')
```

#### Transitivity

What was transitivity again? "The number of transitive triads divided by
all possible transitive triads" @SNASS. So try the transitivity
function.

```{r}
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero"))
```

So, select a part of the network, in this case a focus on node 2.

```{r}
gmatrix1 <- gmatrix
diag(gmatrix1) <- 1
nodes <- gmatrix1[2,] == 1
sel <- gmatrix1[nodes,nodes]
diag(sel) <- 0 
gmatrix2 <- graph_from_adjacency_matrix(sel, mode = "undirected")
plot(gmatrix2)
nodes
```

#### Betweenness

```{r}
igraph::betweenness(g, directed = FALSE)

```

So a lot of ties go through node 1.

#### Dyad census

```{r}
igraph::dyad.census(g)
```

So, 78 mutual dyads, 0 asymmetrical. So, here you can see that this
entails an undirected network. 483 pairs with *no* connections.

```{r}
igraph::triad.census(g)
```

So, this shows that there are 3971 '003' tryads (i.e. no ties), 393
'201' tryads (i.e. A \<-\> B \<-\> C) and 45 '300' triads (complete
graphs)

```{r}
igraph::transitivity(g, type = "global")
```

#### Network visualization

```{r}
# changing V
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 10  #changing the node size based on betweenness of the node!!
V(g)$color = 'green'
plot(g, mode = "undirected")
```

But, we want it more clear

```{r}
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)
```

Hmm, still not what we want.

```{r}
l  #let us take a look at the coordinates
l[34, 1] <- -3.5
plot(g, layout = l)
```

Better

### Twittersphere in the Dutch HoP
```{r}
load("/Users/jannevanheesch/Documents/R. SN/twitter_20190919.RData")
str(twitter_20190919, 1) # see structure to make possible to save in different objects
keyf <- twitter_20190919[[1]]
mydata <- twitter_20190919[[2]]
seats <- twitter_20190919[[3]]


fnet <- mydata$depvars$fnet
atmnet <- mydata$depvars$atmnet
rtnet <- mydata$depvars$rtnet

vrouw <- mydata$cCovars$vrouw
partij <- mydata$cCovars$partij
ethminz <- mydata$cCovars$ethminz
lft <- mydata$cCovars$lft

# if you construct an object for RSiena, covariates are mean centered by default. I would like to
# have the original values again.
ethminz <- ethminz + attributes(ethminz)$mean
partij <- partij + attributes(partij)$mean
vrouw <- vrouw + attributes(vrouw)$mean
lft <- lft + attributes(lft)$mean

str(fnet)

fnet1 <- fnet[, , 1]  #friendship network wave 1
atmnet1 <- atmnet[, , 1]  #atmention network wave 1
# we will use wave 2 and 3 later.
atmnet2 <- atmnet[, , 2]
atmnet3 <- atmnet[, , 3]

table(fnet1, useNA='always') 
fnet1[fnet1 == 10] <- 0
table(fnet1, useNA='always') # this is to assign old 10's to 0's (NA)

atmnet1[atmnet1 == 10] <- 0
atmnet2[atmnet2 == 10] <- 0
atmnet3[atmnet3 == 10] <- 0

```
So, now we use this data to plot and visualize

```{r}
G1 <- igraph::graph_from_adjacency_matrix(atmnet1, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)
plot(G1)

edges <- igraph::as_data_frame(G1, what = "edges")
edges

# the first variable of the data we can attach needs to be some id, thus reorder columns of keyf
keyf <- cbind(keyf$EGOid, keyf[, names(keyf) != "EGOid"])
# the name has been changed as well. Lets correct this
names(keyf)[1] <- "EGOid"

# rebuild the graph.
G1 <- graph_from_data_frame(edges, directed = TRUE, vertices = keyf)

# let us examine the attributes
vertex_attr(G1)

# thus to find the names of our MPs we could now do this:
V(G1)$Naam

#so lets try
plot(G1
)
edge_density(G1)
```

### From directed to reciprocated
```{r}
# define undirected network
atmnet1_un <- atmnet1 == 1 & t(atmnet1) == 1

G2 <- graph_from_adjacency_matrix(atmnet1_un, mode = "undirected", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)

# attach data if you want
edges <- igraph::as_data_frame(G2, what = "edges")
G2 <- graph_from_data_frame(edges, directed = FALSE, vertices = keyf)
plot(G2)
```

### Select some nodes to plot
```{r}
#first make sure we don't end up with MPS who only mention themselves
diag(atmnet1_un) <- 0

# lets find the noisolates
noisolates <- rowSums(atmnet1_un, na.rm = T) > 0
noisolates
# length(noisolates) sum(noisolates) if you select, select both correct nomination network as ego
# characteristics
atmnet1_un_sel <- atmnet1_un[noisolates, noisolates]
# if you are going to use the dataset keyf to add characteristics to the plot later, make sure to
# run the correct selection as well!!!
keyf_sel <- keyf[noisolates, ]

G2_sel <- graph_from_adjacency_matrix(atmnet1_un_sel, mode = "undirected", weighted = NULL, diag = TRUE,
    add.colnames = NA, add.rownames = NA)
G2_sel <- simplify(G2_sel)
plot(G2_sel, mode = "undirected")
```

```{r}
# changing V (node) based on degree
V(G2)$size = igraph::degree(G2) * 1.05  #naturally, you may use a different node-level structural characteristic here. 
plot(G2, mode = "undirected")
```

Now let's play with colors
```{r}
V(G2)$label = as.character(V(G2)$Naam2) #label based on name
V(G2)$label.cex = 1
V(G2)$color <- ifelse(V(G2)$Geslacht == "vrouw", "red", "green") #if woman, then node=red, otherwise green
plot(G2, mode = "undirected")
```

Things like arrow size, curvature, etc
```{r}
E(G2)$arrow.size = 0.4
E(G2)$curved = 0.3
plot(G2, mode = "undirected")
```

```{r}
{
plot.igraph(G2, margin = 0, mode = "udirected")
legend(x = -1, y = -0.75, c("Female", "Male"), pch = 21, col = "#777777", pt.bg = c("red", "green"), pt.cex = 2, cex = 0.8, bty = "n", ncol = 1)
}
```

### Coordinates
```{r}
plot(keyf$X, keyf$Y, xlim = c(-18, 18), ylim = c(-18, 18), col = keyf$Partij_col, pch = 16)
# change node size a bit
V(G2)$size
V(G2)$size = igraph::degree(G2) * 1.05 + 6

# remove the labels
V(G2)$label = ""

# less curvature
E(G2)$curved = 0.1
plot(G2)

owncoords <- cbind(keyf$X, keyf$Y)
owncoords <- owncoords/8
owncoords[, 1] <- (owncoords[, 1] - mean(owncoords[, 1]))
owncoords[, 2] <- (owncoords[, 2] - mean(owncoords[, 2]))
plot.igraph(G2, mode = "undirected", layout = owncoords, rescale = F, margin = c(0, 0, 0, 0), xlim = c(min(owncoords[,
    1]), max(owncoords[, 1])), ylim = c(min(owncoords[, 2]), max(owncoords[, 2])))
```




