---
title: " week 2"
author: "by: René van Son"
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: 72
---

# Revised Research Question & Introduction

Collaboration in the sciences is generally regarded as beneficial, both
for the quality of the research and for the development of scientific
theories and methodologies (@haines2011). Interdisciplinary research, as
a kind of collaboration, has become a focus in many universities, where
curricula are adapted to better connect different disciplines in order
for empirical phenomena to be approached from different perspectives and
to create knowledge integration across disciplines. In fact, some
scholars have argued that research-standards in some disciplines have
already become increasingly interdisciplinary 'below the surface',
pointing at the need for interdisciplinary proposals when trying to
apply for research grants @locatelli2021collective. Others, however,
have argued that while *mentions* of interdisciplinarity (which have
indeed increased!) is reflected by neither a notable increase in
interdisciplinary *work*, nor in the *success* of interdisciplinary
studies (in terms of citations, higher ranking journals, impact, etc.)
@barthel2017interdisciplinary.

One way to better understand interdisciplinarity - its causes,
attributes and consequences - is by taking a social network perspective.
One of the main reasons for taking a social network perspective is
because it treats interdependencies (of observations, persons, etc.) as
theoretically interesting, rather than as violations of assumptions.
Treating researcher ties (in terms of collaborations) as informative may
yield different results than focusing solely on the properties of the
researcher him/herself (e.g. is the researcher him/herself educated
across multiple disciplines). In other words, not only the attributes of
the researcher may have causal properties for the degree of
interdisciplinarity, but the "interdisciplinary-ness" is also dependent
on the *network* of a researcher.

With regards to interdisciplinary collaborations, a social network
analysis will shed light on, firstly, the way in which interdisciplinary
research changed the structure of collaboration networks over time and,
secondly, what the changing structure of these networks means for
researchers who engaged in interdisciplinary research. Following this
line of thinking, we can develop the following research questions:

RQ1: did collaboration networks become more interdisciplinary over time,
i.e. did the structure of the collaboration network change over time
with regards to the frequency/proportion of interdisciplinary
collaborations? RQ2: to what extent does increasing interdisciplinarity
at the macro-level(=network) lead to selective/partial clustering at the
micro level?       To answer these research questions, one needs a
particular kind of data, i.e. data that is relational. For the purpose
here, collaboration data from multiple research departments of Radboud
University will be used.

# Practice

We can describe a network by referring to the diads and the triads. If
they have theoretical meaning, we could ask questions about them. E.g.
we expect 210-configurations than 201-configuration. E.g.: Friendships
are stable relations (usually). These configurations have theoretical
meaning.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#install.packages("igraph") if nec.
require(igraph)
```

Let's try an adjacency matrix to practice: <br>

```{r, echo=FALSE}
matrix1 <- matrix(c(0,0,0,0, #here there are only 0's, there are no (directed) relations 
                   0,0,0,0,
                   0,0,0,0,
                   0,0,0,0),nrow = 4, ncol = 4,byrow = TRUE)
matrix1
```

Above, there are no relations... <br> <br>

```{r, echo=FALSE}
matrix2 <- matrix(c(0,1,0,0, #here there is 1 directed relations, from 1 to 2 (but not from 2 to 1)
                   0,0,0,0,
                   0,0,0,0,
                   0,0,0,0),nrow = 4, ncol = 4,byrow = TRUE)

matrix2
```

Now, there is one (directed) relation: from '1' to '2' (but not from '2'
to '1'!) <br> <br>

```{r, echo=FALSE}
graph1 <- graph_from_adjacency_matrix(
  matrix1)

graph2 <- graph_from_adjacency_matrix(
  matrix2)

graph_matrix1 <- triad_census(graph1)
graph_matrix2 <- triad_census(graph2)
graph_matrix1
graph_matrix2
```

Above, we see the triad counts for matrix1 and matrix2, respectively.
Note that in the first matrix there are four '003' relationships,
whereas in the second matrix there are two '003' relationships and two
'012' relationships. This means: Matrix1: four triads with 0 ties (3
isolates). Matrix2: two triads with 0 ties (3 isolates), two triads with
one directed tie and one isolate. <br> <br> How do these triads come
about? E.g. transitive triads (friends of friends become friends).
Different pathways have different theoretical mechanisms.

Another network characteristic is density (number of ties divided by
possibile ties). Degree centrality is another descriptive charateristic.
How many degrees you have? But in large networks, of course you will
have many relations (out-degrees). So, we normalize it, e.g. max-min
normalization.

Closeness centrality is another one, how central is a node ? Betweenness
centrality is another one. These are all descriptives at the ego-level

Try:

```{r}
#install.packages("RSiena")
require(RSiena)
s501
?s501
?RSiena
dim(s501)

smaller_matrix <- s501[15:25,15:25]
graph_small <- graph_from_adjacency_matrix(smaller_matrix)
plot(graph_small)

```

Okay, so now we have to get centrality measures etc. First degree
centrality

```{r}
smaller_matrix
graph_out <- degree(graph_small, mode = "out")
graph_in <- degree(graph_small, mode = "in")
graph_in
graph_out
```

Closeness

```{r}
closeness(graph_small)
```

```{r}
Betweenness <- betweenness(graph_small)
Betweenness
?betweenness
```

Note that, for example, for V3
